---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

&nbsp;
&nbsp;

This package provides tools to construct sparse Phenotype-Phenotype Network (PPN) from GWAS summary statistics. It enables researchers to infer conditional dependencies among phenotypes by estimating sparse precision matrices using regularized methods. This package is particularly useful for identifying phenotype clusters,visualization of the network structure and multiple phenotype association tests. 

## Installation

You can install the development version of sGGM from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("MeghSubedi/sGGM")
```

## Example

One built-in example dataset is included with the package to facilitate exploration of the sGGM framework. This dataset contains 10,000 rows and 40 columns, representing GWAS summary statistics for 10,000 SNPs across 40 phenotypes.

```{r example}
# load the library
library(sGGM)

# load the example data 
data("example_data")
dim(example_data)
```

```{r}
example_data[1:10,1:8]
```
Lets explore the cluster structure obtained from the correlation matrix and partial correlation matrix. Get_Clusters() function will return the list containing cluster structure from correlation matrix as well as from partial correlation matrix. 

```{r cars}
# Detect clusters from correlation and partial correlation matrix 
Modules<-Get_Clusters(example_data)

# The cluster structure from the correlation matrix 
Modules$B_cor
```
```{r pressure}
# The cluster structure from partial correlation matrix 
Modules$B_Pcor
```

Now, we can use these network modules to test the association of SNPs with multiple phenotype. 

```{r}
# Association test without considering Network modules 
Pval<-Test_Without_Network(example_data,"Wald")
# This will return vector of length 10000 equivalent to the number of SNPs 
length(Pval)
```
```{r}
head(Pval)
```

```{r}
# Association test considering network modules from correlation matrix 
Pval<-Test_With_Network(example_data,Modules$B_cor,method="Wald")
head(Pval)
```
```{r}
dim(Pval)
```

The result is a data frame containing p-values from association tests conducted on phenotypes within each module. These p-values can then be combined to obtain a final p-value for the global association test.

```{r}
# Association test considering network modules from partial correlation matrix 
Pval<-Test_With_Network(example_data,Modules$B_Pcor,"Wald")
head(Pval)
```


